---
title: "BA64036_Assignment1"
author: "Ruthvick Bulagakula"
date: "2023-09-29"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The purpose of this assignment is to set up and use the tools for this course.

## Instructions

1. Install the ISLR library using the install.packages() command. Call the library using the
library(ISLR) command to ensure that the library is correctly installed (10% of total points)


**Calling Installed Library**

```{r}

# If i use install.packages() command rmd file is failing to knit. 
# so, i've removed install.packages() command

library(ISLR) # Using Library to call installed ISLR library

```

2. Create a new R-Notebook (.Rmd) file. In the first code chunk, call the ISLR library and then print the summary of the Carseats dataset. How many observations (rows) this dataset contains? (15% of total points)

**Using Careseats Dataset**

```{r}

summary(Carseats) # Using Summary function to get descriptive statistics of Carseats dataset

```

**No.of Observations in carseats dataset**

```{r}

nrow(Carseats) # Using nrow function to get no of rows in Carsearts dataset


```
3. Using the summary statistics shown above, what is maximum value of the advertising attribute? (15% of total points)

**Maximum Values of the advertising**

```{r}

max(Carseats$Advertising) # Using max function to get maximum value of advertising

```
4. Calculate the IQR of the Price attribute. (15% of total points)

**IQR for Price**

```{r}

IQR(Carseats$Price) # Using IQR function to get Interquartile Range for Price

```
5. Plot the Sales against Price. What do you see in there? Calculate the correlation of the two attributes. What does the sign of the correlation coefficient suggest? (15% of total points)

**Plotting Sales againt Price**

```{r}

# Using plot function to get scatterplot for Price againt Sales
plot(Carseats$Price, Carseats$Sales, xlab = "Price", ylab = "Sales", main = "Price VS Sales", col = "Blue") 

```

**Correlation coefficient for price and sales**

```{r}

cor(Carseats$Price, Carseats$Sales) # Using cor function to Price and Sales correlation

```


